High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2021-06-24 paper/main.tex

Given a timeseries of $n$\textsuperscript{th}-order correlations, we
then need to work ``backwards'' in order to generate the order-zero
timeseries.  If the $n$\textsuperscript{th}-order correlation matrix
at a given timepoint is $m_n$, then we can generate an order $n-1$
correlation matrix (for $n > 1$) by taking a draw from
$\mathcal{N}\left(0, m_n\right)$ and reshaping the resulting vector to
have square dimensions.  Intuitively, the re-shaped matrix will look
like a noisy version of the template matrix, $m_{n-1}$.  (When
$n = 1$, no re-shaping is needed; the resulting $K$-dimensional vector
may be used as the observation at the given timepoint.)  After
independently drawing each timepoint's order $n-1$ correlation matrix
from that timepoint's order $n$ correlation matrix, this process can
be applied repeatedly until $n = 0$.  This results in a
$K$-dimensional timeseries of $T$ observations containing the
specified high-order correlations at orders 1 through $n$.  Following
our approach to generating synthetic data exhibiting known first-order
correlations, we constructed a total of 400 additional multivariate
timeseries, collectively reflecting a total of 4 qualitatively
different patterns of dynamic correlations (i.e., 100 datasets
reflecting each type of dynamic pattern: constant, random, ramping,
and event).  Each timeseries comprised 10 zero-order features (dimensions)
that varied over 300 timepoints.  After applying our dynamic
correlation estimation procedure, this yielded a 100-dimensional
timeseries of first-order
features that could then be used to estimate dynamic second-order
correlations.  (We chose to use $K=10$ zero-order features for our
higher order simulations in order to put the accuracy computations
displayed in Figs.~\ref{fig:synthetic} and~\ref{fig:higher_order_sims}
on a roughly even footing.)

>>>>

Given a timeseries of $n$\textsuperscript{th}-order correlations, we
then need to work ``backwards'' in order to generate the order-zero
timeseries.  If the $n$\textsuperscript{th}-order correlation matrix
at a given timepoint is $m_n$, then we can generate an order $n-1$
correlation matrix (for $n > 1$) by taking a draw from
$\mathcal{N}\left(0, m_n\right)$ and reshaping the resulting vector to
have square dimensions.  To force the resulting matrix to be
symmetric, we remove its lower triangle, and replace the lower
triangle with (a reflected version
of) its upper triangle.  Intuitively, the resulting re-shaped matrix will look
like a noisy (but symmetric) version of the template matrix, $m_{n-1}$.  (When
$n = 1$, no re-shaping is needed; the resulting $K$-dimensional vector
may be used as the observation at the given timepoint.)  After
independently drawing each timepoint's order $n-1$ correlation matrix
from that timepoint's order $n$ correlation matrix, this process can
be applied repeatedly until $n = 0$.  This results in a
$K$-dimensional timeseries of $T$ observations containing the
specified high-order correlations at orders 1 through $n$.  Following
our approach to generating synthetic data exhibiting known first-order
correlations, we constructed a total of 400 additional multivariate
timeseries, collectively reflecting a total of 4 qualitatively
different patterns of dynamic correlations (i.e., 100 datasets
reflecting each type of dynamic pattern: constant, random, ramping,
and event).  Each timeseries comprised 10 zero-order features (dimensions)
that varied over 300 timepoints.  After applying our dynamic
correlation estimation procedure, this yielded a 100-dimensional
timeseries of first-order
features that could then be used to estimate dynamic second-order
correlations.  (We chose to use $K=10$ zero-order features for our
higher order simulations in order to put the accuracy computations
displayed in Figs.~\ref{fig:synthetic} and~\ref{fig:higher_order_sims}
on a roughly even footing.)
